\thetitle

### 2013-09-29 tfpie/tfpie.tex
done

\item from an interaction-oriented language to a compiled,
  batch-oriented language.  
\end{itemize}

@@

We contribute an alternative approach to the second semester that
overcomes these problems and provides a gradual introduction 
to object-oriented programming.  Our approach
starts the second semester by introducing \emph{only} the
concept of programming with objects, while all other aspects of course
remain where they were left off in the previous semester.  This
 allows other concepts to be introduced at the point at
which they are relevant and motivated.  Despite this more gradual
approach, the course accomplishes the goal of reaching industrial
language competence by the end of the semester, covering a super-set
of the concepts and topics covered in the \emph{How to Design Classes}-based
course.

@@

More significantly, Northeastern distinctively emphasizes experiential
education, with almost all Computer Science majors participating in a
6 month ``co-op'' internship after their third semester.  These co-ops
take place at a wide variety of companies, and while most students do
some software development, there is almost no uniformity beyond that.
In particular, there is wide variation in the languages students use
while on co-op. This combination sets the constraints under which we
designed our approach.

@@

@@

Thanks to the the preparation in the first half of the course, we can
cover OO programming in a functional subset of Java in a just a few
lectures.  We then increase the subset of the language we use to
encompass mutation, loops, and mutable data structures and introduce
the underlying design principles. We present \texttt{ArrayList}s,
followed briefly by arrays. Students use, and then implement, hash
tables as well as other mutable and immutable data structures.
Conventional input and output are treated only very briefly, as we
focus instead on both fundamentals and exercises making use of real
APIs such as hashing functions or Twitter posting.  Finally,
\texttt{while}, \texttt{for}, and for-each loops are presented,
following the methodology of \htdc which connects loops to stylized
use of recursive functions with accumulators, a technique the students
now have used for two semesters.

>>>>

\item from an interaction-oriented language and tool suite to a
  compiled, batch-oriented language.
\end{itemize}

@@

We contribute an alternative approach to the second semester that
overcomes these problems and provides a gradual introduction 
to object-oriented programming.  Our approach
starts the second semester by introducing \emph{only} the
concept of programming with objects, while all other aspects of the course
remain where they were left off in the previous semester.  This
 allows other concepts to be introduced at the point at
which they are relevant and motivated.  Despite this more gradual
approach, the course accomplishes the goal of reaching industrial
language competence by the end of the semester, covering a super-set
of the concepts and topics covered in the \emph{How to Design Classes}-based
course.

@@

More significantly, Northeastern distinctively emphasizes experiential
education, with almost all Computer Science majors participating in a
6 month ``co-op'' internship after their third semester.  These co-ops
take place at a wide variety of companies, and while most students do
some software development, there is no single platform or
set of tools that commands majority use.
In particular, there is wide variation in the languages students use
while on co-op. This combination sets the constraints under which we
designed our approach.

@@

We chose the half-way point as the time for transition based on
experience with earlier versions of this course. In particular, we
found that a later transition, while allowing us to present additional
concepts in a controlled environment, did not give students sufficient
time and experience with Java.  Subsequent classes found that students
were strong on fundamentals but weak on Java practice.  The other
alternative, transitioning earlier, would not provide sufficient time
to cover the fundamental topics before the transition.

@@

Thanks to the preparation in the first half of the course, we can
cover OO programming in a functional subset of Java in a just a few
lectures.  We then increase the subset of the language we use to
encompass mutation, loops, and mutable data structures and introduce
the underlying design principles. We present \texttt{ArrayList}s,
followed briefly by arrays. Students use, and then implement, hash
tables as well as other mutable and immutable data structures.
Conventional input and output are treated only very briefly, as we
focus instead on both fundamentals and exercises making use of real
APIs such as hashing functions or Twitter posting.  Finally,
\texttt{while}, \texttt{for}, and for-each loops are presented,
following the methodology of \htdc which connects loops to stylized
use of recursive functions with accumulators, a technique the students
now have used for two semesters.
